---
title: "Resp.Calculations.ST"
author: "Shelly Trigg"
date: "9/30/2019"
output: rmarkdown::github_document
---

Load libraries
```{r}
## load libraries
library(readxl)
library(devtools)
library(dplyr)
library(ggplot2)
library(arm)
```

Read in and format size and weight data
```{r}
Size.Info <- read_xlsx("~/Documents/GitHub/P_generosa/amb_v_varlowpH_juvis/data/Shell_length_data.xlsx") #read Size.info data
#format date column
Size.Info$Date <- format(as.POSIXct(strptime(Size.Info$Date, format ="%Y%m%d")),format="%m.%d.%y")
```

Read in and format cumulative respirometry data
```{r}
# call the cumulative resp table of Lolin raw outputs
cumulative_resp_table <- read.csv(file="~/Documents/GitHub/P_generosa/amb_v_varlowpH_juvis/data/SDR/Cumulative_resp_alpha0.4.csv", header=T, stringsAsFactors = FALSE) #read Size.info data
#format date column
cumulative_resp_table$Date <- format(as.POSIXct(cumulative_resp_table$Date,format="%m.%d.%y"),format="%m.%d.%y")
```

Merge size and weight data with respirometry data
```{r}
x <- merge(cumulative_resp_table, Size.Info, by=c("Date","SDR_position", "Run"))
```

**Analyze blanks** 

```{r}
#subset respirometry data for blanks data
blanks_total <- x[which(x$Type == "blank"),]
#create ID column to distinguish individual data points
blanks_total$ID <- paste(blanks_total$Date,blanks_total$Run, blanks_total$SDR_position, sep = ".")
#plot distributions of blank data points and see if any are outliers
ggplot(blanks_total, aes(x = Date, y = Lpc, color = as.factor(Run))) + geom_point() + theme_bw() + ggtitle("Raw resp. rate in blanks") + ylab("Raw respiration rate (Lpc slope)")
#see if removing outliers makes a tighter distribution
#create data frame with mean and standard deviations for each model
test <- blanks_total %>% group_by(Date,Run,Sw.Condition) %>% summarise(n = n(), mean.lpc = mean(Lpc), mean.leq = mean(Leq), mean.lz = mean(Lz), sd.lpc = sd(Lpc), sd.leq = sd(Leq), sd.lz = sd(Lz))
#review any groups with large standard deviation
test
## runs with SD > 0.1 : 09.05 run 1,run 2, 09.06 run1 var.low, 09.19 run 1 var.low
#now check plots from Resp.LoLin.ST.R to see if models are weird or if trend is abnormal
#exclude blanks that seem weird
blanks_total_filtered <- blanks_total[which(blanks_total$ID != "09.05.19.1.C5" & blanks_total$ID != "09.05.19.1.B5" & blanks_total$ID != "09.05.19.2.B5" & blanks_total$ID != "09.06.19.1.A6" & blanks_total$ID != "09.19.19.1.A1"),]
#create data frame with mean and standard deviations for each model without weird data points
avg.blank <- blanks_total_filtered %>% group_by(Date,Run) %>% summarise(n = n(), mean.lpc = mean(Lpc), mean.leq = mean(Leq), mean.lz = mean(Lz), sd.lpc = sd(Lpc), sd.leq = sd(Leq), sd.lz = sd(Lz))
#review blank averages 
avg.blank
#this resolved the SDs > 0.1
```

**Calculate Respiration Rate**
```{r}
#----------------------------Respiration rate calculation -------------------------------------------

#merge respirometry sample data with average blank data
CALC.table <- merge(avg.blank[,c("Date","Run", "mean.lpc")],x, by=c("Date","Run")) 

#remove lines with NA
CALC.table <- CALC.table[which(CALC.table$Wet.weight.g != "NA"),]

#convert length and wet weight fields to numeric
CALC.table$Length <- as.numeric(CALC.table$Length)
CALC.table$Wet.weight.g <- as.numeric(CALC.table$Wet.weight.g)
CALC.table$Width <- as.numeric(CALC.table$Width)
# NOTE: the raw data is in umol L-1 4ml-1
# "resp.RAW.µgLhr" calculation = ((((((abs(Lpc)) - (BLANK.mean.Lpc))*(4/1000))*(60))*31.998))
# (1) corrects for blank (2) converts to Liters (3) converts to hours (4) converts moles to grams; final unit = µg O2 L-1 h-1


for(i in 1:nrow(CALC.table)) {
  CALC.table$resp.RAW.µgLhr[i] <-((((((abs(CALC.table[i,"Lpc"])) - (CALC.table[i,"mean.lpc"]))*(4/1000))*(60))*31.998))
  CALC.table$resp.shell[i] <- CALC.table$resp.RAW.µgLhr[i]/CALC.table[i,"Length"] #per cm shell length
  CALC.table$resp.weight[i] <- CALC.table$resp.RAW.µgLhr[i]/CALC.table[i,"Wet.weight.g"] #per gram wet weight
  CALC.table$resp.weightXsize[i] <- CALC.table$resp.RAW.µgLhr[i]/(CALC.table[i,"Wet.weight.g"]* CALC.table[i,"Length"]*CALC.table[i,"Width"])#per gram wet weight*shell length
}


#remove wells with poorly modeled data
CALC.table$uniqID <- paste(CALC.table$Date,CALC.table$Run, CALC.table$SDR_position, sep = "_")
CALC.table <- CALC.table[which(CALC.table$uniqID != "09.19.19_1_A6" & CALC.table$uniqID != "09.19.19_1_B3" & CALC.table$uniqID != "09.19.19_1_B4" & CALC.table$uniqID != "09.19.19_1_C3" & CALC.table$uniqID != "10.03.19_1_C4" & CALC.table$uniqID != "10.18.19_1_A6" & CALC.table$uniqID != "10.18.19_1_B4" & CALC.table$uniqID != "10.18.19_2_A3" & CALC.table$uniqID != "10.18.19_3_D1"& CALC.table$uniqID != "10.18.19_3_A1"),]

```

**checking normalization**
```{r}
#determine if normalizing by wet weight or length is better

#create a vector for oct 18 data
resp.data1018 <- CALC.table[which(CALC.table$Date == "10.18.19"),"resp.RAW.µgLhr"]

#calc. correlation
cor(resp.data1018,CALC.table[which(CALC.table$Date == "10.18.19"),"Wet.weight.g"])
#plot 
ggplot(CALC.table[which(CALC.table$Date == "10.18.19"),],aes(x = Wet.weight.g, y = resp.RAW.µgLhr, group = hist.curr.treat, color = hist.curr.treat)) + geom_point()


#create a vector for oc 3 data
resp.data1003 <- CALC.table[which(CALC.table$Date == "10.03.19"),"resp.RAW.µgLhr"]

cor(resp.data1003,CALC.table[which(CALC.table$Date == "10.03.19"),"Wet.weight.g"])
#plot
ggplot(CALC.table[which(CALC.table$Date == "10.03.19"),],aes(x = Wet.weight.g, y = resp.RAW.µgLhr, group = hist.curr.treat, color = hist.curr.treat)) + geom_point()

#create a vector for sept. 19 data
resp.data919 <- CALC.table[which(CALC.table$Date == "09.19.19"),"resp.RAW.µgLhr"]

cor(resp.data919,CALC.table[which(CALC.table$Date == "09.19.19"),"Wet.weight.g"])

#plot

ggplot(CALC.table[which(CALC.table$Date == "09.19.19"),],aes(x = Wet.weight.g, y = resp.RAW.µgLhr, group = hist.curr.treat, color = hist.curr.treat)) + geom_point()

#see if it's one particular group
cor(CALC.table[which(CALC.table$Date == "09.19.19" & CALC.table$hist.curr.treat == "amb_amb"),"resp.RAW.µgLhr" ],CALC.table[which(CALC.table$Date == "09.19.19"& CALC.table$hist.curr.treat == "amb_amb"),"Wet.weight.g"])

cor(CALC.table[which(CALC.table$Date == "09.19.19" & CALC.table$hist.curr.treat == "amb_var.low"),"resp.RAW.µgLhr" ],CALC.table[which(CALC.table$Date == "09.19.19"& CALC.table$hist.curr.treat == "amb_var.low"),"Wet.weight.g"])


cor(CALC.table[which(CALC.table$Date == "09.19.19" & CALC.table$hist.curr.treat == "var.low_amb"),"resp.RAW.µgLhr" ],CALC.table[which(CALC.table$Date == "09.19.19"& CALC.table$hist.curr.treat == "var.low_amb"),"Wet.weight.g"])

cor(CALC.table[which(CALC.table$Date == "09.19.19" & CALC.table$hist.curr.treat == "var.low_var.low"),"resp.RAW.µgLhr" ],CALC.table[which(CALC.table$Date == "09.19.19"& CALC.table$hist.curr.treat == "var.low_var.low"),"Wet.weight.g"])


#It seems ambient parental history groups have weird correlations; mainly amb_var.low. but maybe this is ok? it seems small animals are showing high rates. And it seems real for the most part when checking the plate positions and the LoLin models
```


Write out calculated respirometry rates for blanks and for samples
```{r}
#----------------------OUTPUT - save plots and cumulative tables-----------------------------------------

write.csv(blanks_total,"~/Documents/GitHub/P_generosa/amb_v_varlowpH_juvis/data/SDR/All.blank.resp.rates.csv", row.names=FALSE, quote = FALSE)
write.csv(CALC.table,"~/Documents/GitHub/P_generosa/amb_v_varlowpH_juvis/data/SDR/Final.resp.rates.csv", row.names=FALSE, quote = FALSE) 
```

**plot distribution of raw data per group**
```{r}
ggplot(CALC.table) + geom_density(aes(resp.RAW.µgLhr, group = hist.curr.treat, color = hist.curr.treat)) + ggtitle("Distribution of raw resp. data")
```

**plot distribution of shell length per group**
```{r}
ggplot(CALC.table) + geom_density(aes(Length, group = hist.curr.treat, color = hist.curr.treat)) + ggtitle("Distribution of shell length")
```

**plot distribution of weight per group**
```{r}
ggplot(CALC.table) + geom_density(aes(Wet.weight.g, group = hist.curr.treat, color = hist.curr.treat))+ ggtitle("Distribution of weight")
```

**plot distribution of shell length * weight per group**
```{r}
ggplot(CALC.table) + geom_density(aes(Wet.weight.g*Length, group = hist.curr.treat, color = hist.curr.treat))+ ggtitle("Distribution of weight *shell length")
```


**plot distribution of resp norm'd by weight per group**
```{r}
ggplot(CALC.table) + geom_density(aes(resp.weight, group = hist.curr.treat, color = hist.curr.treat))+ ggtitle("Distribution of resp.norm.to.weight")
```




**Boxplots non-normalized data**
```{r}
#create a group category
CALC.table$hist.curr.treat <- paste(CALC.table$Parental.history,CALC.table$Sw.Condition, sep = "_")

ggplot(CALC.table, aes(x = Parental.history,y = resp.RAW.µgLhr, fill = Sw.Condition)) + geom_boxplot() +facet_wrap(~Date) + theme_bw() + ggtitle("Boxplots of non-normalized respirometry data") + ylab("ug O2/L*hr")

```

**Violin plots with boxplot insets and jitter points for non-normalized data**
```{r}
ggplot(CALC.table, aes(x = hist.curr.treat,y = resp.RAW.µgLhr, fill = Sw.Condition)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.15) + geom_jitter(shape =16, size = 2,position= position_jitter(0.05)) +facet_wrap(~Date) + theme_bw() + ggtitle("Non-normalized respirometry data") + ylab("ug O2/L*hr") + xlab("parental_current treatment") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**anova on non-normalized data on Day 0 (Sept. 5 2019, before treatment started)**
```{r}
#run anova for non-normalized data
#for Sept. 5 trial
a <- aov(resp.RAW.µgLhr ~ Parental.history, data = CALC.table[which(CALC.table$Date == "09.05.19"),])
summary(a)
```

**anova on non-normalized data on Day 1 (Sept. 6 2019)**
```{r}
#for Sept. 6 trial
b <- aov(resp.RAW.µgLhr ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.06.19"),])
summary(b)
```

**anova on non-normalized data on Day 14 (Sept. 19 2019)**
```{r}
#for Sept. 19 trial 
c <- aov(resp.RAW.µgLhr ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.19.19"),])
summary(c)
```


**Boxplots of shell length-normalized data **
```{r}

ggplot(CALC.table, aes(x = Parental.history,y = resp.shell, fill = Sw.Condition)) + geom_boxplot() +facet_wrap(~Date) + theme_bw() + ggtitle("Shell length-normalized respirometry data") + ylab("ug O2/L*hr*cm.shell.length") 
```


**Violin plots with boxplot insets and jitter points for shell length-normalized data**
```{r}
ggplot(CALC.table, aes(x = hist.curr.treat,y = resp.shell, fill = Sw.Condition)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.15) + geom_jitter(shape =16, size = 2,position= position_jitter(0.05)) +facet_wrap(~Date) + theme_bw() + ggtitle("Respirometry data normalized by shell length") + ylab("ug O2/L*hr*cm.shell.length") + xlab("parental_current treatment") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**anova on shell length-normalized data on Day 0 (Sept. 5 2019, before treatment started)**
```{r}
#run anova for non-normalized data
#for Sept. 5 trial
a <- aov(resp.shell ~ Parental.history, data = CALC.table[which(CALC.table$Date == "09.05.19"),])
summary(a)
```

**anova on shell length-normalized data on Day 1 (Sept. 6 2019)**
```{r}
#for Sept. 6 trial
b <- aov(resp.shell ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.06.19"),])
summary(b)
```

**anova on shell length-normalized data on Day 14 (Sept. 19 2019)**
```{r}
#for Sept. 19 trial 
c <- aov(resp.shell ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.19.19"),])
summary(c)
```

**Boxplots of wet weight-normalized data **
```{r}
ggplot(CALC.table, aes(x = Parental.history,y = resp.weight, fill = Sw.Condition)) + geom_boxplot() +facet_wrap(~Date) + theme_bw() + ggtitle("Wet weight-normalized respirometry data") + ylab("ug O2/L*hr*gram.wet.weight") 
```

**Violin plots with boxplot insets and jitter points for wet weight-normalized data**
```{r}
ggplot(CALC.table, aes(x = hist.curr.treat,y = resp.weight, fill = Sw.Condition)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.15) + geom_jitter(shape =16, size = 1,position= position_jitter(0.05)) +facet_wrap(~Date, scale = "free") + theme_bw() + ggtitle("Respirometry data normalized by wet weight") + ylab("ug O2/L*hr*gram.wet.weight") + xlab("parental_current treatment") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**jitter points for wet weight-normalized data**
```{r}
ggplot(CALC.table, aes(x = hist.curr.treat,y = resp.weight, color = hist.curr.treat)) + geom_jitter(shape =16, size = 1,position= position_jitter(0.05)) +facet_wrap(~Date, scale = "free") + theme_bw() + ggtitle("Respirometry data normalized by wet weight") + ylab("ug O2/L*hr*gram.wet.weight") + xlab("parental_current treatment") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**dot plots for wet weight-normalized resp data over time**
```{r}
jpeg("MetabolicRateOverTimeNormToWetWeight.jpg", width = 6,height = 4, units = "in", res = 300)
ggplot(CALC.table, aes(x = Date,y = resp.weight, color = hist.curr.treat)) + geom_point(position = position_dodge(width = 0.5)) + theme_bw() + ggtitle("Metabolic rate normalized by wet weight") + ylab("ug O2/L*hr*gram.wet.weight") + xlab("Date") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
dev.off()
```

**boxplots for wet weight-normalized resp data over time**
```{r}
ggplot(CALC.table, aes(x = Date,y = resp.weight, fill = hist.curr.treat)) + geom_boxplot() + theme_bw() + ggtitle("Respirometry data normalized by wet weight") + ylab("ug O2/L*hr*gram.wet.weight") + geom_point(position = position_dodge(width = 0.75)) + xlab("Date") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**aov of wet weight-normalized resp data over time**

```{r}
#run anova for non-normalized data
#for Sept. 5 trial
resp.wt.3way.aov <- aov(resp.weight ~ Sw.Condition*Parental.history*Date, data = CALC.table[which(CALC.table$Date!="09.05.19"),])
summary(resp.wt.aov)
resp.wt.tuk <- TukeyHSD(aov(resp.weight ~ Sw.Condition*Parental.history, data = CALC.table))
summary(resp.wt.tuk)
```

**glm**
```{r}
#run glm for 10-18
resp.glm <- glm(resp.weight ~ Parental.history*Sw.Condition, data = CALC.table[which(CALC.table$Date=="10.18.19"),])

pchisq(c(114559-91439,55-52,lower.tail=FALSE))

```


**anova on wet weight-normalized data on Day 0 (Sept. 5 2019, before treatment started)**
```{r}
#run anova for non-normalized data
#for Sept. 5 trial
a <- wilcox.test(resp.weight ~ Parental.history, data = CALC.table[which(CALC.table$Date == "09.05.19"),])
summary(a)
```

**anova on wet weight-normalized data on Day 1 (Sept. 6 2019)**
```{r}
#for Sept. 6 trial
b <- aov(resp.weight ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.06.19"),])
summary(b)
```

**anova on wet weight-normalized data on Day 14 (Sept. 19 2019)**
```{r}
#for Sept. 19 trial 
c <- aov(resp.weight ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "09.19.19"),])
summary(c)
TukeyHSD(c)
```


**anova on wet weight-normalized data on Day 14 (Oct. 3 2019)**
```{r}
#for Sept. 19 trial 
d <- aov(resp.weight ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "10.03.19"),])
summary(d)
```

**anova on wet weight-normalized data on Day 14 (Oct. 18 2019)**
```{r}
#for Sept. 19 trial 
e <- aov(resp.weight ~ Sw.Condition*Parental.history, data = CALC.table[which(CALC.table$Date == "10.18.19"),])
summary(e)
TukeyHSD(e)
```

plot distribution of respirometry measurments
```{r}
ggplot(CALC.table, aes(x = resp.weight, color = hist.curr.treat)) + geom_density() + theme_bw() + ggtitle("Respirometry data normalized by wet weight") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
it is poisson distributed so run poisson glm on data

```{r}
g <- glm(resp.weight ~ Parental.history + Sw.Condition + Date, CALC.table, family = poisson(link = "log"))

summary(g)
```
all effects are significant, but the deviance is reduced significantly and the residual deviance is much greater than the degress of freedom so overdispersion exists (https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/). 

Try to correct for this using quasi-poisson model:
```{r}
h <- glm(resp.weight ~ Parental.history + Sw.Condition +Parental.history + Date, CALC.table, family = quasipoisson(link = "log"))

summary(h)

```


compare the models
```{r}
coef1 <- coef(g)
coef2 <- coef(h)
se.coef1 <- se.coef(g)
se.coef2 <- se.coef(h)
models.both <- cbind(coef1, se.coef1, coef2, se.coef2, exponent = exp(coef1))
```


Plotting shell length and wet weight over time
```{r}
#convert size data to df
Size.Info <- data.frame(Size.Info)
#convert classes to numeric
Size.Info$Wet.weight.g <- as.numeric(Size.Info$Wet.weight.g)
Size.Info$Length <- as.numeric(Size.Info$Length)
#remove lines with NA 
Size.Info_noNA <- Size.Info[which(Size.Info$Parental.history!= "NA"),]

#plot length x time boxplots
ggplot(Size.Info_noNA, aes(x = Parental.history, y = Length, fill = Sw.Condition)) + geom_boxplot() + facet_wrap(~Date) + theme_bw() + ggtitle("Shell length over time") + ylab("shell length (cm)")

ggplot(Size.Info_noNA, aes(x = Date, y = Length, color = Parental.history,fill = Sw.Condition)) + geom_boxplot() + theme_bw() + ggtitle("Shell length over time") + ylab("shell length (cm)")

#plot wet.weight x time boxplots
ggplot(Size.Info_noNA, aes(x = Date, y = Wet.weight.g, color = Parental.history,fill = Sw.Condition)) + geom_boxplot() + theme_bw() + ggtitle("Wet weight over time") + ylab("wet weight (g)")

#get averages for each group and date, then plot 

#create a group variable for combining parental history and sw.condition information
Size.Info_noNA$group <- paste(Size.Info_noNA$Parental.history,Size.Info_noNA$Sw.Condition, sep = "_")

#create a data frame that contains data grouped by date, parental history, and sw. condition
test <- group_by(Size.Info_noNA, Date, Parental.history, Sw.Condition)

#For shell length, calculate average, standard deviation, and standard error
test_mean <- summarize(test, LengthAvg = mean(Length), LengthSD = sd(Length), LengthSE = sd(Length)/(sqrt(n())))
test_mean$group <- paste(test_mean$Parental.history, test_mean$Sw.Condition, sep = "_")

#set dodge position
pd <- position_dodge(0.4)

#Plot shell length average over time as dots with error bars
ggplot(test_mean, aes(x = Date, y = LengthAvg, color = group, group = group)) + geom_point(position = pd) + geom_errorbar(data = test_mean, aes(ymin = LengthAvg-LengthSE, ymax = LengthAvg+LengthSE), width = 0.1, position = pd, linetype = "dashed") + geom_line(position = pd) + theme_bw()+ ylab("mean shell length (cm)") +labs(color = "history_treatment", title ="mean shell length over time (error bars = SE)" )

#For wet weight, calculate average, standard deviation, and standard error
test_Wtmean <- summarize(test, WtAvg = mean(Wet.weight.g, na.rm = TRUE), WtSD = sd(Wet.weight.g, na.rm = TRUE), WtSE = sd(Wet.weight.g, na.rm = TRUE)/(sqrt(n())))

#For wet weight data, create a group variable for combining parental history and sw.condition information
test_Wtmean$group <- paste(test_Wtmean$Parental.history, test_Wtmean$Sw.Condition, sep = "_")

#Plot wet weight average over time as dots with error bars
ggplot(test_Wtmean, aes(x = Date, y = WtAvg, color = group, group = group)) + geom_point(position = pd) + geom_errorbar(data = test_Wtmean, aes(ymin = WtAvg-WtSE, ymax = WtAvg+WtSE), width = 0.1, position = pd, linetype = "dashed") + geom_line(position = pd) + theme_bw() + ylab("mean wet weight (g)") + labs(color = "history_treatment", title ="mean wet weight over time (error bars = SE)" )
```
