---
title: "2016 pteropod survival proportions analysis"
author: "Shelly Trigg, modified by Ariana Huffmyer"
date: "12/13/2023"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

This script analyzes brood survival. 

# Set up 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, load_libraries, echo = FALSE}
library(readxl)
library(plyr)
library(ggplot2)
library(FSA)
library(survminer)
library(survival)
library(lme4)
library(lmerTest)
library(knitr)
library(coxme)
library(survMisc)
library(gridExtra)
library(tidyverse)
```

# Load data 

```{r}
#survival data
df <- read_xlsx("data/Brood_Survival/Tank5_6_Broodstock_survival_201811-20190317.xlsx")

df$Date <- as.POSIXct(strptime(df$Date, format = "%Y%m%d"))

str(df$Date)
```

Calculate proportion survival at each time point. 
```{r}
df<-df%>%
  group_by(treatment, tank)%>%
  mutate(prop_surv=live/first(live))
```

Change days relative conditioning to be + # after the conditioning period has ended. 
```{r}
df<-df%>%
  mutate(factor_DaysConditioning=as.factor(DaysRelativeToConditioning))%>%
  mutate(factor_DaysConditioning=if_else(factor_DaysConditioning=="41", "+7", 
                                            if_else(factor_DaysConditioning=="50", "+16", 
                                                    if_else(factor_DaysConditioning=="68", "+34", 
                                                            if_else(factor_DaysConditioning=="77", "+43", factor_DaysConditioning)))))%>%
  mutate(factor_DaysConditioning=as.factor(factor_DaysConditioning))

df$factor_DaysConditioning<-factor(df$factor_DaysConditioning, levels=c("-8", "-1", "6", "13", "23", "33", "+7", "+16", "+34", "+43"))
```

Plot survival by time. We only have n=1 replicate per treatment (1 tank per treatment). Shade the conditioning period beginning 20181129 and ending 20190115. 

```{r}
library(ggh4x)

shading_start <- 2.15
shading_end <- 6.15

surv_plot<-df%>%
  
  ggplot(data=., aes(x=interaction(Date, factor_DaysConditioning), group=treatment, y=prop_surv, colour=treatment))+
  
  geom_rect(
    xmin = shading_start, xmax = shading_end,
    ymin = -Inf, ymax = Inf, fill = "lightgray", alpha = 0.2, 
    colour="lightgray"
  ) +
  
  geom_line(linewidth=2)+
  
  scale_colour_manual(values=c("lightblue", "orange"), name="Conditioning Treatment", labels=c("Ambient pH", "Variable Low pH"))+
  ylab("Proportion Survival")+
  xlab("Date \nDays Relative to Conditioning")+
  geom_text(label="Conditioning Period", x=4, y=1, colour="black", size=4)+
  ylim(0,1)+
  ggtitle("Broodstock Survival")+
  
  scale_x_discrete(guide = "axis_nested") +
  theme_classic()+
  theme(
    text=element_text(size=12, colour="black"),
    axis.title=element_text(size=12, colour="black", face="bold"), 
    legend.title=element_text(size=12, colour="black", face="bold"), 
    legend.position="top"
  ); surv_plot

ggsave(filename="figures/brood_survival/broodstock_survival.png", surv_plot, width=10, height=6)
```


We do not have any replicates of the survival for this experiment from what I can tell, so we will not proceed with survival analysis. Keeping the below section for reference. 


I also could not find any reference to STACKED_d df or columns including MOATS or Jars in the data or in previous versions of this analysis file. 

# Survival analysis using Kaplan Meier *NOT USING, FOR REFERENCE ONLY* 

Create a long form data frame
```{r}
STACKED_d<-df%>%pivot_longer()
```


```{r}
#create a new blank dataset for data to input into survival analysis
nMoats <- nlevels(STACKED_d$MOATS)
nJars <- nlevels(STACKED_d$Jar)
ds <- data.frame(Day = rep(0, (nMoats*nJars)))
ds$Dead <- 0
ds$Moats <- NA
ds$pHtreat <- NA
ds$DOtreat <- NA
ds$Treat <- NA
#View(ds)

#loop to create a dataframe in the format to make a "surv" object for survival analysis
#the format records the day of events (deaths) and the day the pteropod leave the experiment (either by going missing or the end of the experiment)
#days are relative to the start of the experiment (i.e. days of exposure; collection day = 0)
index <- 1
for(i in 1:nMoats){
  for(j in 1:nJars){
    dc <- subset(STACKED_d, MOATS == levels(STACKED_d$MOATS)[i] & Jar == levels(STACKED_d$Jar)[j])
    dc <- dc[order(dc$Date),]
    if(dc$Status[1] == "Dead"){
      ds$Day[index] <- 1
      ds$Dead[index] <- 1
    }
    nDays <- length(dc$Date)
    for(k in 2:nDays){
      if(dc$Status[k-1] == "Live" && dc$Status[k] == "Dead"){
        ds$Day[index] <- k
        ds$Dead[index] <- 1
      }
      if(dc$Status[k-1] == "Live" && dc$Status[k] == "Missing"){
          ds$Day[index] <- k-1
          ds$Dead[index] <- 0
      }
    }
    if(dc$Status[nDays] == "Live"){
      ds$Day[index] <- nDays
      ds$Dead[index] <- 0
    }
    ds$Moats[index] <- as.character(dc$MOATS[1])
    ds$pHtreat[index] <- dc$pH[1]
    ds$DOtreat[index] <- dc$DO[1]
    ds$Treat[index] <- dc$treatment[1]
    index <- index + 1
  }
}
levels(factor(ds$Moats))
#View(ds)
str(ds)

#make treatment variables factors
ds$Treat <- factor(ds$Treat)
ds$pHtreat <- factor(ds$pHtreat)
ds$DOtreat <- factor(ds$DOtreat)


#make a surv object to pass into coxme

surv <- Surv(time = ds$Day, event = ds$Dead, type = "right")

```


surival plots of simple K-M surivial probabilites; models does in include interaction design or random MOATS term
```{r, echo = FALSE, eval = FALSE}

jpeg("~/Documents/GitHub/pteropod_pHxDO_metabolomics/survival/2016_pteropod_survival_paul_code_adapted_files/surv_curv.jpg", width = 1800, height = 800, res = 200)

ggsurvplot(survfit(surv ~ Treat, ds), risk.table = FALSE, pval = FALSE, conf.int = TRUE, palette = "npg", break.x.by = 1, xlim = c(1,9) , xlab = "Time (days)") #pH and DO treatments

dev.off()



```

cox test
```{r}


survFit1 <- coxph(surv ~ pHtreat + DOtreat + pHtreat*DOtreat, ds )
summary(survFit1)

survFitME <- coxme(surv ~ pHtreat*DOtreat + (1|Moats), ds)
summary(survFitME)

survFitME1 <- coxme(surv ~ Treat + (1|Moats), ds)
summary(survFitME1)

#r-squared
#https://rdrr.io/cran/survMisc/man/rsq.html
rsq(survFit1, sigD=2)

#pairwise comparisons

library(multcomp)

glht(model = survFitME1, linfct = mcp(Treat = "Dunnett"),alternative = "greater")
```


The standard deviation of the MOATS random effect is very small (0.009) which suggests that MOATS does not really affect the outcome. Also the liklihood ratio test (Chisq) value is the same (13.05), meaning that including MOATS as a random effect does not improve the model. We likely do not need to include it as a random effect. 

**Describing the cox model**
I based the interpretation off of this article: http://www.sthda.com/english/wiki/cox-proportional-hazards-model

1. Statistical significance. Z = the Wald statistic which evaluates whether the beta coefficient (impact of the covariate) is statistically significantly different from 0. We can conclude that low pH and low DO did not have statistically significant coefficients but their interaction did (p = 0.00117).

2. Regression coefficient ("coef"). A coef < 0 means hazard is lower and > 0 means hazard is higher. So, low pH or low DO treatment led to a lower risk of death (AKA greater survival), but the interaction of low pH and low DO led to greater risk of death (AKA less survival). 

3. Hazard ratio ("exp(coef)"): the effect size of the covariates. Low pH treatment reduced the risk of death by a factor of 0.6697 or ~33%, and low DO treatment reduced the risk of death by a factor of 0.5942 or ~41%. Low pH and low DO treatment increased the risk of death by a factor of 3.3422 or 330%. 

4. confidence intervals of the hazard ratios. So low pH factor reduced risk of death is 0.6697 + 

5. Global statistical significance of the model: p = 0.004528, based on the liklihood ratio test (but also the same p-value when logrank test or wald test is used)

6. Rsquared: coefficient of determination < 0.2 means predictive power is low. Coefficient of determination is not generalized to censored data, so measure of explained randomness or measure of explained variation are more popular. These values are still < 0.2 in this case, so the model has weak predictive power on survival. Reference (https://apha.confex.com/apha/134am/techprogram/paper_135906.htm)
