---
title: "Juvenile Suvival Analysis (Oct 18)"
author: "Shelly Trigg, modified by Ariana Huffmyer"
date: "12/13/2023"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

This script analyzes juvenile survival.

# Set up 

Load libraries
```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
library(readxl)
library(cowplot)
library(emmeans)
```

# Survival data from counting (Steven)

Read in data 
```{r}
counts<-read_xlsx("data/juv_survival/20191018_survival.xlsx")%>%select(!`percent.survival(init. 98animals/silo)`)
metadata<-read_xlsx("data/juv_survival/juv_metadata.xlsx")%>%rename(silo=Tank.ID, parental.history=Parental.history, sw.condition=Sw.Condition)

counts$starting.animals<-98
```

Calculate percent survival as number of alive animals/number of initial animals. 
```{r}
counts$prop_surv<-counts$avg.count/counts$starting.animals
```

Create a code for treatment. 
```{r}
counts$code<-paste0(counts$parental.history, "-", counts$sw.condition) 
```

# Survival data from images 

```{r}
fileName <- "data/juv_shell_length/20191018_ShellLength.xlsx"
#code from: https://stackoverflow.com/questions/12945687/read-all-worksheets-in-an-excel-workbook-into-an-r-list-with-data-frames
#define the function for reading in the excel file with multiple tabs
read_excel_allsheets <- function(fileName, tibble = FALSE) {
  sheets <- readxl::excel_sheets(fileName)
  x <- lapply(sheets, function(X) readxl::read_excel(fileName, sheet = X))
  if(!tibble) x <- lapply(x, as.data.frame)
  names(x) <- sheets
  x
}
# create R object for excel file 
sheets <- read_excel_allsheets("data/juv_shell_length/20191018_ShellLength.xlsx")
list2env(sheets ,.GlobalEnv)
#create data frame from excel datafile
dt <- data.frame()
for(i in 1:length(sheets)){
  dtemp <- data.frame(sheets[i])
  dtemp$Image <- names(sheets)[i]
  colnames(dtemp) <- c("Area","Mean", "Min", "Max", "Angle", "Length (inch)", "Length (cm)", "Image")
  dt <- rbind(dt, dtemp)
}

# add Silo Information 
dt$Silo <- gsub(".*_","",dt$Image)

#Read in the parental history and treatment conditions
meta_data <- read_excel("data/juv_survival/juv_metadata.xlsx")

#change column name to match that in dt
colnames(meta_data)[1] <- "Silo"

#merge dt data with meta data
dt <- merge(dt,meta_data, by = "Silo")

#create a new column with parental history and current treatment info
dt$code <- paste(dt$Parental.history, dt$Sw.Condition,sep = "-")

dt<-dt%>%
  rename(parental.history=Parental.history, sw.condition=Sw.Condition, silo=Silo)

meta_data<-meta_data%>%
  rename(silo=Silo, parental.history=Parental.history, sw.condition=Sw.Condition)

head(dt)
```

Create data frame with final counts from images
```{r}
img_counts <- data.frame(table(dt$silo))
colnames(img_counts) <- c("silo", "live.animals")
img_counts <- merge(img_counts, meta_data, by = "silo")
img_counts$code <- paste0(img_counts$parental.history, "-", img_counts$sw.condition)
```

Add starting number of animals
```{r}
img_counts$starting.animals<-98
```

Calculate proportion survival
```{r}
img_counts$prop_surv<-img_counts$live.animals/img_counts$starting.animals
```

# Plot survival as reaction norms

Plot survival from count data

```{r}
surv_plot1<-counts%>%
  
  ggplot(aes(x=sw.condition, group=interaction(sw.condition, parental.history), y=prop_surv, colour=parental.history))+
  
  geom_boxplot(linewidth=1, outlier.shape = NA)+
  geom_point(size=3, position=position_jitterdodge(0.1))+
  
  scale_colour_manual(values=c("lightblue", "orange"), name="Parental Treatment", labels=c("Ambient pH", "Variable Low pH"))+
  ylab("Proportion Survival")+
  xlab("Juvenile Condition")+
  scale_x_discrete(labels=c("Ambient pH", "Variable Low pH"))+
  ylim(0,1.5)+
  ggtitle("Juvenile Survival - Counts")+
  
  theme_classic()+
  theme(
    text=element_text(size=12, colour="black"),
    axis.title=element_text(size=12, colour="black", face="bold"), 
    legend.title=element_text(size=12, colour="black", face="bold"), 
    legend.position="none"
  ); surv_plot1

surv_plot1a<-counts%>%
  group_by(sw.condition, parental.history)%>%
  summarize(mean=mean(prop_surv, na.rm=TRUE), se=sd(prop_surv, na.rm=TRUE)/sqrt(length(prop_surv)))%>%
  
  ggplot(aes(x=sw.condition, group=interaction(sw.condition, parental.history), y=mean, colour=parental.history))+
  
  geom_point(size=3, position=position_dodge(0.1))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), position=position_dodge(0.1), width=0)+
  geom_line(aes(group=parental.history), size=1, position=position_dodge(0.1))+
  
  scale_colour_manual(values=c("lightblue", "orange"), name="Parental Treatment", labels=c("Ambient pH", "Variable Low pH"))+
  ylab("Proportion Survival")+
  xlab("Juvenile Condition")+
  scale_x_discrete(labels=c("Ambient pH", "Variable Low pH"))+
  ylim(0,1.5)+
  ggtitle("Juvenile Survival - Counts")+
  
  theme_classic()+
  theme(
    text=element_text(size=12, colour="black"),
    axis.title=element_text(size=12, colour="black", face="bold"), 
    legend.title=element_text(size=12, colour="black", face="bold"), 
    legend.position="right"
  ); surv_plot1a

surv_grid<-plot_grid(surv_plot1, surv_plot1a, ncol=2, nrow=1, rel_widths=c(0.8,1))

ggsave(filename="figures/juv_survival/juv_survival_counts.png", surv_grid, width=10, height=6)
```

Plot survival from image data 
```{r}
surv_plot2<-img_counts%>%
  
  ggplot(aes(x=sw.condition, group=interaction(sw.condition, parental.history), y=prop_surv, colour=parental.history))+
  
  geom_boxplot(linewidth=1, outlier.shape = NA)+
  geom_point(size=3, position=position_jitterdodge(0.1))+
  
  scale_colour_manual(values=c("lightblue", "orange"), name="Parental Treatment", labels=c("Ambient pH", "Variable Low pH"))+
  ylab("Proportion Survival")+
  xlab("Juvenile Condition")+
  scale_x_discrete(labels=c("Ambient pH", "Variable Low pH"))+
  ylim(0,1.5)+
  ggtitle("Juvenile Survival - Images")+
  
  theme_classic()+
  theme(
    text=element_text(size=12, colour="black"),
    axis.title=element_text(size=12, colour="black", face="bold"), 
    legend.title=element_text(size=12, colour="black", face="bold"), 
    legend.position="none"
  ); surv_plot2

surv_plot2a<-img_counts%>%
  group_by(sw.condition, parental.history)%>%
  summarize(mean=mean(prop_surv, na.rm=TRUE), se=sd(prop_surv, na.rm=TRUE)/sqrt(length(prop_surv)))%>%
  
  ggplot(aes(x=sw.condition, group=interaction(sw.condition, parental.history), y=mean, colour=parental.history))+
  
  geom_point(size=3, position=position_dodge(0.1))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), position=position_dodge(0.1), width=0)+
  geom_line(aes(group=parental.history), size=1, position=position_dodge(0.1))+
  
  scale_colour_manual(values=c("lightblue", "orange"), name="Parental Treatment", labels=c("Ambient pH", "Variable Low pH"))+
  ylab("Proportion Survival")+
  xlab("Juvenile Condition")+
  scale_x_discrete(labels=c("Ambient pH", "Variable Low pH"))+
  ylim(0,1.5)+
  ggtitle("Juvenile Survival - Images")+
  
  theme_classic()+
  theme(
    text=element_text(size=12, colour="black"),
    axis.title=element_text(size=12, colour="black", face="bold"), 
    legend.title=element_text(size=12, colour="black", face="bold"), 
    legend.position="right"
  ); surv_plot2a

surv_grid2<-plot_grid(surv_plot2, surv_plot2a, ncol=2, nrow=1, rel_widths=c(0.8,1))

ggsave(filename="figures/juv_survival/juv_survival_images.png", surv_grid2, width=10, height=6)
```

# ANOVA on survival 

Run ANOVA on count data 

```{r}
model1<- aov(prop_surv ~ parental.history * sw.condition, data = counts)
summary(model1)

emm<-emmeans(model1, ~parental.history)
pairs(emm)

emm<-emmeans(model1, ~sw.condition)
pairs(emm)

emm<-emmeans(model1, ~parental.history | sw.condition)
pairs(emm)
```
Parental condition and juvenile condition are significant in count data. Posthoc shows no difference in parental treatment at ambient pH juvenile treatment. But the variable low pH parental treatment are lower under variable low pH juvenile treatment. Overall, parental ambient treatment is higher than exposed parent treatment. Ambient juvenile treatment is higher than exposed juvenile treatment. 

Run ANOVA on image data 

```{r}
model2<- aov(prop_surv ~ parental.history * sw.condition, data = img_counts)
summary(model2)

emm<-emmeans(model2, ~parental.history)
pairs(emm)

emm<-emmeans(model2, ~sw.condition)
pairs(emm)
```
Ambient parental condition is higher than exposed. Ambient juvenile condition is also higher than exposed. 
